\acknowledgements{\OnePageChapter{}

  Reflecting back on my time in Boulder, it is almost overwhelming to realize how
  many people made my time here so special. I am so lucky to have had an amazing
  support system that I could always rely on. Thank you to everyone.

  It's safe to say that I would not be writing this if it were not for my
  advisors Meredith Betterton and Matt Glaser.  Meredith, I am grateful for your
  guidance and everything you taught me over the years. Your broad knowledge
  base in physics, math, and biology as well as your ability to communicate
  science always impressed me.  Matt, you have always been someone I could
  count on for support.  You are a great scientist and friend, and I
  always admired your skepticism and insight. While technically a collaborator,
  I would like to thank my unofficial third advisor, Loren Hough.  Some of our
  scientific discussions were the most rewarding I've had. I'd also like to
  thank collaborator and committee member Franck Vernerey for our fruitful
  discussions.  I appreciate your insight and help on projects, and I regret
  that I did not get to work with you more.  Since I've hit almost my whole
  committee, it would be a shame to leave out Joe MacLennan.  Joe, I always
  enjoyed our chats in the hallway.  Thanks for being my biggest mohawk fan
  and for the coloring ideas. Shout out as well to others in the department:
  Tom DeGrand, Mike Hermele, Paul Romatschke, Leo Radzihovsky, John Bohn, Mike
  Ritzoller, Mike Dubson, Oliver Nix, Kirsten Apocada, Annett Baumgartner,
  Dakota Nanton, Lauren Moreno, Jeanne Nijhowne, and others for stimulating
  discussions, advice, and help.

  As for the rest of my group, thank you to Chris Edelmaier, Adam Lamson, Robert
  Blackwell, and Zach Gergely for all of the help and advice. Special thanks to
  Laura Maguire and Hui-Shun Kuan for working on these projects with me; it was
  a pleasure. And then, there is Jeffrey Moore. Jeff, you may have had the
  biggest influence on me over the three years since we've known each other. My
  Windows, git-less, and VIM-less days feels like a lifetime ago. Thank you for
  helping me become the computational physicist I am today; I couldn't have done
  it without you.  I will never forget your patience and advice.  You are an
  incredible problem solver and code monkey, and I cannot wait to see what you
  accomplish. I hope I can help on the other side.

  Outside of work, thank you to my loving family for their support.  I am so
  lucky to have such loving parents. From the days of AYSO soccer to my eleventh
  year of college, you have been cheering me along all the way.  It truly means
  a lot.  Lindsay, you've always been the greatest source of emotional support.
  Thanks for always knowing how to make me laugh. I love you all so much.

  I do not think I would have survived graduate school without my friends in
  Boulder. The Bro-op has been one of my greatest sources of joy, adventure, and
  personal connection.  Short-term members Katie, Stacey, and Dimitri: it was a
  blast, and I wish we had more time to live together.  Stephen, Tom, and Matt,
  you guys will forever be my brothers. I am so grateful for our time over the
  years. You helped me keep my sanity. Thank you. Living with one the world's
  most brilliant physicists, electrical engineers, and a polymath was quite the
  experience. You taught me the value of hard work, skepticism, communication,
  and critical thinking---even when it is inconvenient. Tom, thank you for being
  the best friend I could ever hope for.  I'll never forget all you have done
  for me.  I would also like to thank the Dinner Club for wonderful memories
  over the ($>5$!) years, Martha for crafting, Molly for our conversations, and
  Phoebe for the pottery help and laughs.

  Ok, I'm almost done. Many thanks to Boulder's premier party mixed frisbee
  team, All Jeeps All Night, for having me as a practice player. Captains Marek,
  Tristan, and Grant, you made this all possible. Thanks for the time you've
  spent organizing everything; you've brought me and others countless great
  memories. Shout out to Joliene, Stevie, and all the wonderful baristas at
  Pekoe for keeping me fueled with free coffee for three years. It was fun being
  the guy who got free stuff, but no one knew why. Thanks to Stephanie and the
  wonderful PTs for keeping me up and running over the years. To all the awesome
  musicians who played at our house parties, thank you for some of my fondest
  memories.  To my fellow Cocktail Burrito band mates Justin Griffith and Tom
  Nummy (and biggest fan Dani), thanks for letting me play rhythmless guitar and
  bass and sing off-key with you guys.  Finally, thanks to the Armer boys, HMS
  Beagle, and the 41st house for always hosting fun events.  Clearly, I think
  work-life balance is important.  Thanks again everyone who helped me with
  that.

  Finally, science ain't free. Thanks for the support from the SMRC
  (DMR-1420736), Biofrontiers Computing Core (NIH 1S10OD012300), and RMACC
  Summit supercomputer (awards ACI-1532235 and ACI-1532236).
}


