\chapter{Hydrogels yay}\label{ch03}
\section{Pore size}
\section{Porogens}
I tried to make gels with a larger pore size by using porogens.  These are nanoscale particles that can be mixed into the gel precursor solution and then removed/dissolved after crosslinking.  The idea is that then they would leave pores of a better size and also make it easier for the gel to swell to equilibrium.  I tried alginate nanospheres (which could then be removed by chelating the calcium ions) and large molecular-weight dextrans that could be removed by digesting with dextranase.  Neither of these options worked out.  I couldn't get the nanospheres to reliably resuspend in the precusor solution.  The dextran/dextranase system was more promising, but too much dextran remained in the hydrogels even after dextranase digestion.  Also the dextranase chewed up the FSFG.
\subsection{Alginate nanospheres}
Figures:
\begin{enumerate}
\item diagram of fabrication process
\item DLS results on size distribution
\end{enumerate}
\subsection{Dextran/dextranase system}
Figures:
\begin{enumerate}
\item diagram of porogen process + molecular structure of dextran, dextranase?
\item removal of fluorescent dextran from gel
\item comparison of gels with and without 250 kD and 70kD dextran
\item weird mCherry suppression?
\end{enumerate}
\section{Polymerization using confocal microscope}
5)	Writing thin gels with confocal microscope
a.	50 um minimum line thickness – good
b.	Wrote thin lines and donuts
c.	Tried using photoinhibitors to reduce stray crosslinking, but didn’t work well enough
d.	Swelling/buckling was a problem
\section{Nup and transport factor constructs}
6)	Constructs
a.	SSSG as a negative control for FSFG gels (this actually works)
b.	Covalently-tethered NTF2 (didn’t work)
c.	FSFG cct2 (works) and cct3 (doesn’t express)
d.	Kap121 with GFP-NLS (only tested once, didn’t enter gel)
\section{Dye-labeling and free dye}
7)	Fixed free dye problem
a.	Rigorous labeling protocol and testing, keep frozen until use
b.	Tried a few different labeling chemistries (maleimide, NHS and SDP esters)
c.	Made GFP-NTF2 (didn’t bind to FSFG)
